---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quantdr

<!-- badges: start -->
  [![Travis build status](https://travis-ci.com/elianachristou/quantdr.svg?branch=master)](https://travis-ci.com/elianachristou/quantdr)
<!-- badges: end -->

quantdr: Dimension Reduction Techniques for Conditional Quantiles
====================================================================

The R package 'quantdr' performs dimension reduction techniques for conditional quantiles by estimating the fewest linear combinations of X that contain all the information on that function.  For details of the methodology, see [Christou, E. (2020) Central quantile subspace. *Statistics and Computing*, 30, 677–695](https://link.springer.com/article/10.1007/s11222-019-09915-8).

The main function of the package is `cqs`, which estimates the directions of the central quantile subspace.  Once the directions are determined, one can form the new sufficient predictors and estimate the conditional quantile function using `llqr`.  

## Installation

You can install the released version of quantdr from [CRAN](https://CRAN.R-project.org) with:

``` r 
install.packages("quantdr") 
```

and the development version from [GitHub](https://github.com/) with: 

<!-- You can install the development version of quantdr from [GitHub](https://github.com/) with:  -->

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("elianachristou/quantdr")
```
## Example

This is a basic example which shows you how to solve the problem. 

```{r}
library(quantdr)

## basic example code - a homogeneous multi-index model

# Setting
set.seed(1234)
n <- 100; p <- 10; taus <- c(0.1, 0.25, 0.5, 0.75, 0.9)
x <- matrix(rnorm(n * p), n, p); error <- rnorm(n)
y <- 3 * x[, 1] + x[, 2] + error

# true direction that spans each central quantile subspace
beta_true <- c(3, 1, rep(0, p - 2))
beta_true / sqrt(sum(beta_true^2))
dir1 <- x %*% beta_true

# Estimate the directions of each central quantile subspace
out <- matrix(0, p, length(taus))
for (i in 1:length(taus)) {
  out[, i] <- cqs(x, y, tau = taus[i], d = 1, dtau = 1)$qvectors
}
out

# compare the estimated direction with the true one using the angle between the two subspaces
library(pracma)
for (i in 1:length(taus)) {
print(subspace(out[, i], beta_true) / (pi / 2)) # the angle is measured in radians, so divide by pi/2
  }

# Estimate and plot the conditional quantile function using the new sufficient predictors
newx <- x %*% out
par(mfrow=c(2,3))
for (i in 1:length(taus)) {
  plot(dir1, y, xlab = "sufficient direction", ylab = "y", main = taus[i], pch = 16)
  qhat <- llqr(newx[, i], y, tau = taus[i])$ll_est
  points(dir1, qhat, pch = 16, col = "red")
}
```

Another example using the Boston housing data from the `MASS` library in `R`. 

```{r, warning = FALSE}
library(MASS)
attach(Boston)
# read the data 
y <- medv
x <- cbind(rm, log(tax), ptratio, log(lstat))
n <- length(y); p <- dim(x)[2]

# plot the estimated coefficient of each predictor variable for multiple quantiles
tau <- seq(0.1, 0.9, by = 0.005)
beta_hat <- matrix(0, p, length(tau))

for (k in 1:length(tau)) {
  out <- cqs(x, y, tau = tau[k])
  beta_hat[, k] <- out$qvectors[, 1:out$dtau] # the estimated dimension of the central subspace is 1
}
   
par(mfrow=c(2,2))
plot(tau, beta_hat[1, ], type = 'l', xlab = 'Quantile', main = 'RM', ylab = 'Coefficient')
plot(tau, beta_hat[2, ], type = 'l', xlab = 'Quantile', main = 'log(TAX)', ylab = 'Coefficient')
plot(tau, beta_hat[3, ], type = 'l', xlab = 'Quantile', main = 'PTRATIO', ylab = 'Coefficient')
plot(tau, beta_hat[4, ], type = 'l', xlab = 'Quantile', main = 'log(LSTAT)', ylab = 'Coefficient')

```
