---
title: "quantdr: Dimension Reduction for Conditional Quantiles"
author: 
- name: Eliana Christou
  affiliation: University of North Carolina at Charlotte
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{quantdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
abstract: "While dimension reduction techniques have been well developed, they focus on the entire conditional distribution of the data.  There are a lot of research areas where extremes are important and require the study of the conditional quantiles of the response given the predictor variables.  `quantdr` is an `R` package that performs dimension reduction to conditional quantiles by determining the directions that span the central quantile subspace (CQS)."  
keywords: "central quantile subspace, dimension reduction, quantile regression"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Getting Started

To get started you need to first install the package using the command 
```{r, eval=FALSE}
install.packages("quantdr")
```
and then use it during any `R` session using the command
```{r}
library(quantdr)
```
For an overview on the available help files of the package use
```{r, eval=FALSE}
help(package = "quantdr")
```
The package includes the following functions:

* `llqr` Local linear quantile estimation
* `llqrcv` Cross-Validation bandwidth selection for local linear quantile estimation
* `cqs` Central quantile subspace
* `bic_d` Modified-BIC type criterion for structural dimension selection

To get help on a specific function you can type `help(function.name)` or `?function.name` for a convenient shorthand.  Try
```{r, eval=FALSE}
help(cqs)
?cqs
```

To run examples in `R` you can use the examples used to illutrated that the commands are executable by pasting them into the `R` session.

# Introduction
## Introduction

-----

* Make a list
    * with 4 spaces nested
    
1. Ordered list
1. Ordered list

Definition
: say the definition

*italic*
**bold**

```{r, results='hide'}
# if i want to write code
add <- function(a, b) a + b
add(10, 20)
```

```{r, results='asis'}
# if i want to write code
add <- function(a, b) a + b
add(10, 20)
```

Consider submitting the vignette to the Journal of Statistical Software of The R Journal. 

