---
title: "quantdr"
author: 
- name: Eliana Christou
  affiliation: University of North Carolina at Charlotte
output: rmarkdown::html_vignette
header-includes: |
  ```{=latex}
  \newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
  ```
vignette: >
  %\VignetteIndexEntry{quantdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
abstract: "While dimension reduction techniques have been well developed, they focus on the entire conditional distribution of the data.  There are a lot of research areas where extremes are important and require the study of the conditional quantiles of the response given the predictor variables.  `quantdr` is an `R` package that performs dimension reduction to conditional quantiles by determining the directions that span the central quantile subspace (CQS)."  
keywords: "central quantile subspace, dimension reduction, quantile regression"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Getting Started

To get started you need to first install the package using the command 
```{r, eval=FALSE}
install.packages("quantdr")
```
and then use it during any `R` session using the command
```{r}
library(quantdr)
```
For an overview on the available help files of the package use
```{r, eval=FALSE}
help(package = "quantdr")
```
The package includes the following functions:

* `llqr` Local linear quantile estimation
* `llqrcv` Cross-Validation bandwidth selection for local linear quantile estimation
* `cqs` Central quantile subspace
* `bic_d` Modified-BIC type criterion for structural dimension selection

To get help on a specific function you can type `help(function.name)` or `?function.name` for a convenient shorthand.  Try
```{r, eval=FALSE}
help(cqs)
?cqs
```

To run examples in `R` you can use the examples used to illutrated that the commands are executable by pasting them into the `R` session.

# Central quantile subspace
## Overview

Data visualization and non-parametric modeling can become very chalennging when data is *high-dimensional*.  This has led to the development of many dimension reduction techniques, which focus on the entire conditional distribution of the response given the predictor variables.  However, when specific aspects of the conditional distribution are of interest, these methods can provide more directions than necessary.

Quantile regession (QR) has received a lot of attention since its inception from Koenker and Bassett (1978) and it has been an attractive tool for research areas where extremes are important.  Christou (2020) proposed a dimension reduction technique for conditional quantiles that finds the fewest linear combinations of the predictor variable $\mathbf{X}$ necessary for determining the conditional quantile.     

To give further intuition, for a univariate response $Y$ and a $p$ dimensional predictor $\mathbf{X}$, the method focuses on finding a $p \times d_{\tau}$ matrix $\mathbf{B}_{\tau}$, where $\tau \in (0, 1)$ the quantile level and $d_{\tau} \leq p$, such that 
\begin{eqnarray*}
Y \independent Q_{\tau}(Y|\mathbf{X}) | \mathbf{B}_{\tau}^{\top} \mathbf{X},
\end{eqnarray*}
for $Q_{\tau}(Y|\mathbf{X})$ the $\tau$th conditional quantile of $Y$ given $\mathbf{X}$.  The space spanned by $\mathbf{B}_{\tau}$ is called the *$\tau$th quantile dimension reduction subspace* for the regression of $Y$ on $\mathbf{X}$.  The intersection of all $\tau$th dimension reduction subspaces is called the *$\tau$th central quantile subspace* ($\tau$-CQS) and is denoted by $\mathcal{S}_{Q_{\tau}(Y|\mathbf{X})}$.  Therefore, the goal is to identify the directions of $\mathcal{S}_{Q_{\tau}(Y|\mathbf{X})}$ and form the new sufficient predictors.     

The purpose of this paper is to demonstrate the implementation of the functions appearing in the `quantdr` package, along with some examples.  The main function of the package is `cqs`, which returns the estimated directions of the $\tau$-CQS.  

## The algorithm in a glance

Christou (2020) proved that

* $\boldsymbol{\beta}_{\tau} \in \mathcal{S}_{Q_{\tau}(Y|\mathbf{X})}$, where
\begin{eqnarray*}
(\alpha_{\tau}, \boldsymbol{\beta}_{\tau}) = \arg \min_{(a_{\tau}, \mathbf{b}_{\tau})} E \{ Q_{\tau}(Y| \mathbf{A}^{\top} \mathbf{X}) - a_{\tau} - \mathbf{b}_{\tau}^{\top}\mathbf{X}\}
\end{eqnarray*}

* $E\{Q_{\tau} (Y|U_{\tau})\mathbf{X}\} \in \mathcal{S}_{Q_{\tau}(Y|\mathbf{X})}$, where $U_{\tau}$ is a measurable function of $\mathbf{B}_{\tau}^{\top}\mathbf{X}$, provided that $Q_{\tau}(Y|U_{\tau})\mathbf{X}$ is integrable.  

This implies that we can set $\boldsymbol{\beta}_{\tau, 0}=\boldsymbol{\beta}_{\tau}$ as the initial vector and then create more using $\boldsymbol{\beta}_{\tau,j}=E\{Q_{\tau}(Y | \boldsymbol{\beta}_{\tau, j-1}^{\top}\mathbf{X})\mathbf{X}\}$, for $j=1, \dots, p-1$.

-----

* Make a list
    * with 4 spaces nested
    
1. Ordered list
1. Ordered list

Definition
: say the definition

*italic*
**bold**

```{r, results='hide'}
# if i want to write code
add <- function(a, b) a + b
add(10, 20)
```

```{r, results='asis'}
# if i want to write code
add <- function(a, b) a + b
add(10, 20)
```

Consider submitting the vignette to the Journal of Statistical Software of The R Journal. 

